|

### 2024-04-21 manuscript.Rmd
Added autobuild files

---
title: |
  | Neural correlates of the uncanny valley effect
  | for robots and hyper-realistic masks 
  | (for submission to Computers in Human Behavior)
author: |
  | Shona Fitzpatrick, Ailish K. Byrne, Alex Headley, Jet G. Sanders,
  | Helen Petrie, Rob Jenkins & Daniel H. Baker (daniel.baker@york.ac.uk)
date: "University of York"
output:
  pdf_document:
    toc: no
  bookdown::pdf_document2:
    fig_caption: yes
    toc: no
    keep_tex: yes
  word_document: default
  html_document: default
bibliography: references.bib
csl: elife.csl
---

@@

processdata <- 0  # this flag determines the amount of processing, with 4 levels:
# 0 - do no processing, generate the pdf using existing versions of all figures
# 1 - generate figures using the processed group data (requires 120MB of storage)
# 2 - do statistics and bootstrapping using processed data
# 3 - download all raw data and do preprocessing (requires ~60GB of storage)

@@

rawdir <- 'local/raw/'
rawdir <- '/Volumes/USB128/UncannyValley/'
processeddir <- 'local/output/'
# processeddir <- '/Volumes/USB128/UncannyValley/'

if (!dir.exists('local/')){dir.create('local/')}
# if (!dir.exists('local/robotsmvpa/')){dir.create('local/robotsmvpa/')}
# if (!dir.exists('local/masksmvpa/')){dir.create('local/masksmvpa/')}

@@

@@

```{r stimexamples, fig.cap="Illustrative stimuli from the same categories as used in Experiments 1 and 2. Row (a) shows a human face, an android and a robot, all against white backgrounds. Row (b) shows a human face, a hyper-realistic silicone mask, and a carnival mask, all against natural backgrounds. Images were taken from a variety of sources that permit reuse in academic contexts and in most cases were not part of the stimulus set from the experiments. The silicone mask image was taken by the authors, and was used in Experiment 2.", fig.align="center", echo=FALSE, fig.width=8, fig.height=8}

@@

We analysed response data by calculating d-prime scores for each condition, derived from the hit rate and false alarm rate [@Macmillan2005]. For the human conditions, the hit rate was the proportion of human images correctly identified as human, and the false alarm rate was the proportion of robot or mask images that were incorrectly judged as being human. For the robot and mask conditions, the hit rate was the proportion of robot/mask images correctly identified as not being human, and the false alarm rate was the proportion of human images that were incorrectly judged as being non-human (note that this means the false alarm rate was the same for the robot and android conditions, and for the silicone and Halloween mask conditions). We capped infinite d-prime values (which occur e.g. when the hit rate is 1) at an arbitrary ceiling of 5. We log transformed the reaction times (which typically have positive skew) and performed all averaging and statistical analysis on the logarithmic values.

@@

if (file.exists('local/GroupDataRobots.RData')){load('local/GroupDataRobots.RData')}  
if (processdata > 0){

@@

We first explored the behavioural results for identification of human versus non-human stimuli. We calculated d-prime scores to compare sensitivity across conditions. Sensitivity was highest for identifying robots (d' = `r round(conddprimes[1],digits=2)`), but still well above chance for both the human (d' = `r round(conddprimes[3],digits=2)`) and android (d' = `r round(conddprimes[2],digits=2)`) conditions. The Bayes factor score for a one-way ANOVA comparing these three conditions indicated very substantial evidence ($log_{10}BF_{10} =$ `r round(log10(bfdprime),digits=2)`) for a difference between conditions, as illustrated in Figure \ref{fig:RobotsData}a. Pairwise Bayesian t-tests between conditions indicate very convincing differences in sensitivity between robots and androids ($log_{10}BF_{10} =$ `r round(log10(bfcontrasts[1]),digits=2)`) and robots and humans ($log_{10}BF_{10} =$ `r round(log10(bfcontrasts[2]),digits=2)`). The difference between androids and humans ($log_{10}BF_{10} =$ `r round(log10(bfcontrasts[3]),digits=2)`) was also very substantial.

@@

```{r RobotsData, fig.cap="Summary of response data and grand mean ERP for Experiment 1. Panel (a) shows d-prime scores for identifying images of human (blue), android (green) and robot (red) faces. Small points show individual participants, and the larger symbols with error bars indicate the group mean and bootstrapped 95 percent confidence intervals. Panel (b) plots reaction times in the same format (note the logarithmic x-axis). Panel (c) shows the grant mean ERP across all participants and conditions, pooled across electrodes P6, P8, PO6 and PO8 (see inset). The shaded region around the curve illustrates the 95 percent confidence interval, and the grey rectangle at the foot indicates the stimulus duration.", fig.align="center", echo=FALSE, fig.width=12, fig.height=12}

@@

```{r RobotsMVPA, fig.cap="Univariate and multivariate comparisons across image type. Panel (a) shows the ERPs comparing human (blue) and robot (red) face images, and panel (b) compares human (blue) and android (green) faces. Panels (c) and (d) show multivariate pattern classification accuracy for the same comparisons. Points at y = -8 and y = 25 indicate Bayes factor scores for comparisons between ERPs (a,b) and comparing classification accuracy to chance (50 percent correct; c,d).", fig.align="center", echo=FALSE, fig.width=12, fig.height=12}

@@

if (file.exists('local/GroupDataMasks.RData')){load('local/GroupDataMasks.RData')}  
if (processdata > 0){

@@

The results of Experiment 2 were similar to those of Experiment 1, despite using a quite different stimulus set involving images of humans wearing masks, rather than robots. Sensitivity was highest for identifying Halloween masks (d' = `r round(conddprimes[1],digits=2)`), but still well above chance for both the human (d' = `r round(conddprimes[3],digits=2)`) and silicone mask (d' = `r round(conddprimes[2],digits=2)`) conditions. The Bayes factor score for a one-way ANOVA comparing these three conditions indicated very substantial evidence ($log_{10}BF_{10} =$ `r round(log10(bfdprime),digits=2)`) for a difference between conditions, as illustrated in Figure \ref{fig:MasksData}a. Pairwise Bayesian t-tests between conditions indicate very convincing differences in sensitivity between Halloween and silicone masks ($log_{10}BF_{10} =$ `r round(log10(bfcontrasts[1]),digits=2)`) between Halloween masks and humans ($log_{10}BF_{10} =$ `r round(log10(bfcontrasts[2]),digits=2)`), and between silicone masks and humans ($log_{10}BF_{10} =$ `r round(log10(bfcontrasts[3]),digits=2)`). Unlike in Experiment 1, there were no convincing reaction time differences between conditions ($log_{10}BF_{10} =$ `r round(log10(bfRT),digits=2)`), as illustrated in Figure \ref{fig:MasksData}b.

```{r MasksData, fig.cap="Summary of response data and grand mean ERP for Experiment 2. Panel (a) shows d-prime scores for identifying images of human faces (blue), silicone masks (green) and Halloween masks (red). Small points show individual participants, and the larger symbols with error bars indicate the group mean and bootstrapped 95 percent confidence intervals. Panel (b) plots reaction times in the same format (note the logarithmic x-axis). Panel (c) shows the grant mean ERP across all participants and conditions, pooled across electrodes P6, P8, PO6 and PO8 (see inset). The shaded region around the curve illustrates the 95 percent confidence interval, and the grey rectangle at the foot indicates the stimulus duration.", fig.align="center", echo=FALSE, fig.width=12, fig.height=12}

@@

```{r MasksMVPA, fig.cap="Univariate and multivariate comparisons across image type for Experiment 2. Panel (a) shows the ERPs comparing human faces (blue) and Halloween masks (red), and panel (b) compares human faces (blue) and silicone masks (green). Panels (c) and (d) show multivariate pattern classification accuracy for the same comparisons. Points at y = -8 and y = 25 indicate Bayes factor scores for comparisons between ERPs (a,b) and comparing classification accuracy to chance (50 percent correct; c,d).", fig.align="center", echo=FALSE, fig.width=12, fig.height=12}

>>>>

---
title: |
  | Neural correlates of the uncanny valley effect
  | for robots and hyper-realistic masks 
author: |
  | Shona Fitzpatrick, Ailish K. Byrne, Alex Headley, Jet G. Sanders,
  | Helen Petrie, Rob Jenkins & Daniel H. Baker (daniel.baker@york.ac.uk)
date: "University of York"
output:
  bookdown::pdf_document2:
    fig_caption: yes
    toc: no
    keep_tex: yes
  word_document: default
  pdf_document:
    toc: no
  html_document: default
bibliography: references.bib
csl: elife.csl
---

@@

processdata <- 0  # this flag determines the amount of processing, with 3 levels:
# 0 - do no processing, generate the pdf using existing versions of all figures
# 1 - generate figures using the processed group data (requires 120MB of storage)
# 2 - do statistics and bootstrapping using processed data

@@

if (!dir.exists('local/')){dir.create('local/')}
if (!dir.exists('local/robotsmvpa/')){dir.create('local/robotsmvpa/')}
if (!dir.exists('local/masksmvpa/')){dir.create('local/masksmvpa/')}

@@

  if (!file.exists('local/channel.mat')){
  osfproject <- osf_retrieve_node('uzbns')
  osffiles <- osf_ls_files(osfproject)
  fid <- which(osffiles$name=='channel.mat')
  osf_download(osffiles[fid,], path='local/',progress=TRUE)}

  if (!file.exists('local/channeldata.csv')){
  osfproject <- osf_retrieve_node('uzbns')
  osffiles <- osf_ls_files(osfproject)
  fid <- which(osffiles$name=='channeldata.csv')
  osf_download(osffiles[fid,], path='local/',progress=TRUE)}

@@

```{r stimexamples, fig.cap="Illustrative stimuli from the same categories as used in Experiments 1 and 2. Row (a) shows a human face, an android and a robot, all against white backgrounds. Row (b) shows a human face, a hyper-realistic silicone mask, and a carnival mask, all against natural backgrounds. Images were taken from a variety of sources that permit reuse in academic contexts and in most cases were not part of the stimulus set from the experiments. The silicone mask image was taken by the authors, and was used in Experiment 2.", fig.align="center", out.width='80%', echo=FALSE}

@@

We analysed response data by calculating d-prime ($d'$) scores for each condition, derived from the hit rate and false alarm rate [@Macmillan2005]. For the human conditions, the hit rate was the proportion of human images correctly identified as human, and the false alarm rate was the proportion of robot or mask images that were incorrectly judged as being human. For the robot and mask conditions, the hit rate was the proportion of robot/mask images correctly identified as not being human, and the false alarm rate was the proportion of human images that were incorrectly judged as being non-human (note that this means the false alarm rate was the same for the robot and android conditions, and for the silicone and Halloween mask conditions). We capped infinite d-prime values (which occur e.g. when the hit rate is 1) at an arbitrary ceiling of 5. We log transformed the reaction times (which typically have positive skew) and performed all averaging and statistical analysis on the logarithmic values.

@@

  if (!file.exists('local/GroupDataRobots.RData')){
  osfproject <- osf_retrieve_node('uzbns')
  osffiles <- osf_ls_files(osfproject)
  fid <- which(osffiles$name=='GroupDataRobots.RData')
  osf_download(osffiles[fid,], path='local/',progress=TRUE)}

load('local/GroupDataRobots.RData')
if (processdata > 0){

@@

We first explored the behavioural results for identification of human versus non-human stimuli. We calculated d-prime scores to compare sensitivity across conditions. Sensitivity was highest for identifying robots ($d' =$ `r round(conddprimes[1],digits=2)`), but still well above chance for both the human ($d' =$ `r round(conddprimes[3],digits=2)`) and android ($d' =$ `r round(conddprimes[2],digits=2)`) conditions. The Bayes factor score for a one-way ANOVA comparing these three conditions indicated very substantial evidence ($log_{10}BF_{10} =$ `r round(log10(bfdprime),digits=2)`) for a difference between conditions, as illustrated in Figure \ref{fig:RobotsData}a. Pairwise Bayesian t-tests between conditions indicate very convincing differences in sensitivity between robots and androids ($log_{10}BF_{10} =$ `r round(log10(bfcontrasts[1]),digits=2)`) and robots and humans ($log_{10}BF_{10} =$ `r round(log10(bfcontrasts[2]),digits=2)`). The difference between androids and humans ($log_{10}BF_{10} =$ `r round(log10(bfcontrasts[3]),digits=2)`) was also very substantial.

@@

```{r RobotsData, fig.cap="Summary of response data and grand mean ERP for Experiment 1. Panel (a) shows d-prime scores for identifying images of human (blue), android (green) and robot (red) faces. Small points show individual participants, and the larger symbols with error bars indicate the group mean and bootstrapped 95 percent confidence intervals. Panel (b) plots reaction times in the same format (note the logarithmic x-axis). Panel (c) shows the grand mean ERP across all participants and conditions, pooled across electrodes P6, P8, PO6 and PO8 (see inset). The shaded region around the curve illustrates the 95 percent confidence interval, and the grey rectangle at the foot indicates the stimulus duration.", fig.align="center", echo=FALSE}

@@

```{r RobotsMVPA, fig.cap="Univariate and multivariate comparisons across image type. Panel (a) shows the ERPs comparing human (blue) and robot (red) face images, and panel (b) compares human (blue) and android (green) faces. Panels (c) and (d) show multivariate pattern classification accuracy for the same comparisons. Points at y = -8 and y = 25 indicate Bayes factor scores for comparisons between ERPs (a,b) and comparing classification accuracy to chance (50 percent correct; c,d).", fig.align="center", echo=FALSE}

@@

  if (!file.exists('local/GroupDataMasks.RData')){
  osfproject <- osf_retrieve_node('uzbns')
  osffiles <- osf_ls_files(osfproject)
  fid <- which(osffiles$name=='GroupDataMasks.RData')
  osf_download(osffiles[fid,], path='local/',progress=TRUE)}

load('local/GroupDataMasks.RData')

if (processdata > 0){

@@

The results of Experiment 2 were similar to those of Experiment 1, despite using a quite different stimulus set involving images of humans wearing masks, rather than robots. Sensitivity was highest for identifying Halloween masks ($d' =$ `r round(conddprimes[1],digits=2)`), but still well above chance for both the human ($d' =$ `r round(conddprimes[3],digits=2)`) and silicone mask ($d' =$ `r round(conddprimes[2],digits=2)`) conditions. The Bayes factor score for a one-way ANOVA comparing these three conditions indicated very substantial evidence ($log_{10}BF_{10} =$ `r round(log10(bfdprime),digits=2)`) for a difference between conditions, as illustrated in Figure \ref{fig:MasksData}a. Pairwise Bayesian t-tests between conditions indicate very convincing differences in sensitivity between Halloween and silicone masks ($log_{10}BF_{10} =$ `r round(log10(bfcontrasts[1]),digits=2)`) between Halloween masks and humans ($log_{10}BF_{10} =$ `r round(log10(bfcontrasts[2]),digits=2)`), and between silicone masks and humans ($log_{10}BF_{10} =$ `r round(log10(bfcontrasts[3]),digits=2)`). Unlike in Experiment 1, there were no convincing reaction time differences between conditions ($log_{10}BF_{10} =$ `r round(log10(bfRT),digits=2)`), as illustrated in Figure \ref{fig:MasksData}b.

```{r MasksData, fig.cap="Summary of response data and grand mean ERP for Experiment 2. Panel (a) shows d-prime scores for identifying images of human faces (blue), silicone masks (green) and Halloween masks (red). Small points show individual participants, and the larger symbols with error bars indicate the group mean and bootstrapped 95 percent confidence intervals. Panel (b) plots reaction times in the same format (note the logarithmic x-axis). Panel (c) shows the grand mean ERP across all participants and conditions, pooled across electrodes P6, P8, PO6 and PO8 (see inset). The shaded region around the curve illustrates the 95 percent confidence interval, and the grey rectangle at the foot indicates the stimulus duration.", fig.align="center", echo=FALSE}

@@

```{r MasksMVPA, fig.cap="Univariate and multivariate comparisons across image type for Experiment 2. Panel (a) shows the ERPs comparing human faces (blue) and Halloween masks (red), and panel (b) compares human faces (blue) and silicone masks (green). Panels (c) and (d) show multivariate pattern classification accuracy for the same comparisons. Points at y = -8 and y = 25 indicate Bayes factor scores for comparisons between ERPs (a,b) and comparing classification accuracy to chance (50 percent correct; c,d).", fig.align="center", echo=FALSE}
